Merging datasets through deep learning

### 2018-08-23 paper2018/introduction2.tex
new version of intro

In this paper, we propose a novel approach to the problem of joining different surface representations of the same entity, inspired in part by recent advances in deep learning.  Our approach depends on (a) creating a function that maps surface forms into a set of vectors such that alternate forms for the same entity are closest in vector space, (b) indexing these vectors using a space partitioning algorithm to find the forms that can be potentially joined together.  Our approach for merging datasets is much more flexible than prior data driven approaches (e.g., \cite{He:2015:SJS:2824032.2824036}, \cite{auto-join-joining-tables-leveraging-transformations}) because it can be applied to any data provided one can create a mapping function such that joinable surface forms of the entity are closer in vector space than disparate forms.  For instance, in face recognition and person re-identification, there has been a significant research effort aimed at creating deep learning models that learn to map different surface forms of a face closer together in a vector space.  In our proposed approach, we can now join two columns if they have alternate faces for the same entity, whereas in prior work, those faces needed to either have been joined before in some existing dataset, or the transforms of a face need to be discoverable as a program, which is near impossible for faces.  Our approach for merging datasets is also efficient because it eliminates quadratic comparisons of cell values in each column by using a mechanism that is itself data-driven.  Because we use space partitioning algorithms to find surface forms that are potentially joinable, the search process is efficiently eliminates large parts of the vector space from consideration for any given queried vector.  Across all cell values that are considered for a join, the efficiency of the search is $\mathcal{O}(n\log{}n)$.

Specifically, the first part of our solution to the join problem involves building a deep learning model that performs metric learning.  For the closely related problem of face recognition and person re-identification, deep neural networks have been used successfully to build models that map faces of the same entity closer in vector space estimate relative to faces from other entities.  To test the generality of our approach, we tried to build a deep learning model model for names of people and companies. As in face recognition, our system for metric learning is built by training a `triplet siamese network' to learn to map different surface forms of the same entity closer in vector space than different forms of the same entity.  Entity names have very different properties from faces, so we needed to adapt the approaches used in face recognition to the problem of matching entity names with a novel method to choose the sample for training, and novel loss functions as objectives for the metric learning problem.  At training, the network is given a triplet consisting of an `anchor' (e.g. \textit{Douglas Adams}), a positive element (e.g. \textit{Douglas Noel Adams}) and a negative element (e.g., John Adams), and it learns to produce a small distance estimate for the positive pair (close to 0), and a large distance estimate for the negative pair, using an objective function. A classification form of this function that simply learns whether two names belong to the same entity can be used to perform fuzzy joins, but then we would need a blocking strategy analogous to what has been used in the literature to identify the set of joinable pairs \cite{auto-join-joining-tables-leveraging-transformations}.

Our observation is that one can actually exploit what the siamese network produces to eliminate this blocking step altogether.  Specifically, the last hidden layer of the siamese network is effectively a `vector embedding' that contains the critical features needed for computing the distance estimate.  If such models are pre-built for a given entity type, we can then use the models to produce vector embeddings for all cell values in the two columns to be joined in a single linear pass, and index just the second column values with a space partitioning algorithm that lays out vectors in n-dimensional space.  Then, for each cell value in the first column to be joined, querying the k-nearest neighbors should provide the set of values in the second column that can be joined. In our work, we used approximate nearest neighbor algorithms, which have been applied successfully to billions of items \cite{JDH17}.

@@

Our key contributions are as follows:
\begin{itemize}
\item We outline a general approach to the problem of merging datasets for different surface forms of the same entity.  Our approach consists of building a deep neural network model for specific entity types and using a space partitioning algorithm to find the nearest neighbors to join.  To test the generality of our approach, we built triplet loss based siamese networks for ~200,000 people's names and ~70,000 company names respectively, drawn from Wikidata.  We show that such networks did learn to map surface forms of the same entity closer in vector space.  Specifically, if we 83\% and 79\% of variants of the entity name appeared in the top 20 neighbors for each entity.  Moreover, within the 20 neighbors, 77\% and 75\% of the positive names were closer to the anchor than negative names for people and companies respectively.  Our results suggest that deep learning models can be used successfully for data driven joins, assuming that such models are built for common entity types.
\item Training neural networks for deep metric learning is difficult because one cannot show quadratic combinations of the data to the system.  In face recognition, a number of techniques exist for triplet selection, and methods for selecting triples is an active area of research.  We developed a new technique for triplet selection for entity names because the techniques from face recognition do not generalize well to entity names.  We also developed a novel loss function to train the network to discriminate entity names from names of other entities.  We compare the results for this loss function with the ones that have been used for face recognition, and show that this loss function can perform much better than the existing functions for deep learning in this domain.
\item We also perform a complexity analysis of how such models can then be used to perform joins across different surface forms of the same entity efficiently, with a complexity of $\mathcal{O}(n\log{}n)$.  
\end{itemize}

>>>>

In this paper, we propose a different approach to the problem of joining different surface representations of the same entity, inspired in part by recent advances in deep learning.  Our approach depends on (a) creating a function that maps surface forms into a set of vectors such that alternate forms for the same entity are closest in vector space, (b) indexing these vectors using a space partitioning algorithm to find the forms that can be potentially joined together.  Our approach for merging datasets is much more flexible than prior data driven approaches (e.g., \cite{He:2015:SJS:2824032.2824036}, \cite{auto-join-joining-tables-leveraging-transformations}) because it can be applied to any data provided one can create a mapping function such that joinable surface forms of the entity are closer in vector space than disparate forms.  For instance, in face recognition and person re-identification, there has been a significant research effort aimed at creating deep learning models that learn to map different surface forms of a face closer together in a vector space (this technique is often referred to as deep metric learning).  In our proposed approach, we can now join two columns if they have alternate faces for the same entity, whereas in prior work, those faces needed to either have been joined before in some existing dataset, or the transforms of a face need to be discoverable as a program, which is near impossible for faces.  Our approach for merging datasets is also efficient because it eliminates quadratic comparisons of cell values in each column by using a mechanism that is itself data-driven.  Because we use space partitioning algorithms (such as approximate nearest neighbor) to find surface forms that are potentially joinable, the search process is efficiently eliminates large parts of the vector space from consideration for any given queried vector.  Across all cell values that are considered for a join, the efficiency of the search is $\mathcal{O}(n\log{}n)$.  Implementations exist for nearest neighbor algorithms that have been applied to billions of vectors \cite{JDH17}, so the approach is practical for most datasets.

The idea of building joint embeddings for merging followed by nearest neighbors search has been applied recently to a slightly different problem of linking relational tuple embeddings with embeddings of other relational tuples or unstructured text \cite{Bordawekar18}, \cite{IDEL18}).  While our approach is conceptually similar to these two pieces of work, our contribution is that we developed general purpose embedding models for merging alternate surface forms of key entities such as people and companies leveraged from linked open data. Using Wikidata as our ground truth, we built models for datasets with ~1.1M people's names (~200K identities) and 130K company names (~70K identities).  Deep metric learning is known to be a difficult problem in the closely related space of face recognition and person-re-identification \cite{DBLP:conf/cvpr/SchroffKP15}.  To build these models for people and companies, we developed a novel approach for selecting samples for training, as well as developed a novel loss function, as we describe below.

As in face recognition, our system for metric learning is built by training a `triplet siamese network' to learn to map different surface forms of the same entity closer in vector space than different forms of the same entity.  At training, the network is given a triplet consisting of an `anchor' (e.g. \textit{Douglas Adams}), a positive element (e.g. \textit{Douglas Noel Adams}) and a negative element (e.g., John Adams), and it learns to produce a small distance estimate for the positive pair (close to 0), and a large distance estimate for the negative pair, using an objective function. The last hidden layer of the siamese network is effectively a `vector embedding' that contains the critical features needed for computing the distance estimate.  There has been a significant amount of research in the problem of triplet selection, starting with the original Facenet work \cite{DBLP:conf/cvpr/SchroffKP15}.  For any given training set, the key question is how to choose the set of negatives to teach the system the right mapping quickly.  In prior work, for instance, various triplet mining techniques have been used to estimate which negatives might be good to pair with positives for training.  As one example, \cite{DBLP:conf/cvpr/SchroffKP15} perform pairwise comparisons within a batch of training examples to identify \textit{semi-hard negatives} which are negatives that are slightly further away from the anchor than the positive example, but not distant enough from the anchor by a sufficient margin.  One problem with this approach is that it relies on the existence of very large training sets and batches, so that one can find a sufficient number of negative examples for training in a batch.  Our observation is that we can re-use the approximate nearest neighbors approach for triplet selection as well.  Because our ultimate goal is to train the system to learn to map positives closer to the anchor and negatives further away from the anchor, we could select negatives that appear in the nearest neighbor set for input vectors, before any training has taken place.  Depending on the nature of the dataset, these negatives could well be \textit{hard negatives}, i.e., negatives that are closer to the anchor than the positives.  In fact, in our work, most of the triplets we selected had \textit{hard negatives}.  \cite{DBLP:conf/cvpr/SchroffKP15} suggest that using only \textit{hard negatives} might be problematic for metric learning, but in our work, we used this approach successfully to train two separate models for people and companies respectively.  Our observation is that triplet selection based on a nearest neighbors algorithm is a principled and efficient mechanism for selecting triples to train deep metric learning models.

There have also been a number of different objective functions that have been proposed to build better models.  For the problem of building models for people's names, we found that a modification to the objective function proposed in \cite{DBLP:conf/cvpr/SchroffKP15} was better than three other functions that have been proposed in the literature. We show the performance of the different functions for the two datasets we considered.

@@

Our key contributions are as follows:
\begin{itemize}
\item We outline a general approach to the problem of merging datasets for different surface forms of the same entity.  Our approach consists of building a deep neural network model for specific entity types and using a space partitioning algorithm to find the nearest neighbors to join.  To test the generality of our approach, we built triplet loss based siamese networks for 200K people's names and 70K company names respectively, drawn from Wikidata.  We show that such networks did learn to map surface forms of the same entity closer in vector space, with a recall rate of 81\% and 75\% after training for 20 neighbors.  Moreover, within the 20 neighbors, 93\% and 96\% of the positive names were closer to the anchor than negative names for people and companies respectively, suggesting good precision.  Our results suggest that deep learning models can be used successfully for data driven joins, assuming that such models are built for common entity types.
\item Training neural networks for deep metric learning is difficult because one cannot show quadratic combinations of the data to the system.  In face recognition, a number of techniques exist for triplet selection, and methods for selecting triples is an active area of research.  We developed a new technique for triplet selection for entity names based on an approximate nearest neighbors algorithm over input vectors.  We also explored whether a modified loss function can be used to train the network more effectively, and found that it does in fact perform better for training on people's names.
\item A key element of our argument is that once these embedding models have been built, they can be applied to perform joins efficiently for a new dataset.  We test if this is true with previously built face recognition models, because of the paucity of large datasets for companies and people, and the wealth of datasets in the face recognition literature.  For the most part, deep metric models for faces have been used with a classification layer have been used to identify a specific individual, or for clustering faces to measure goodness of the clustering algorithm.  Because we use these existing models for a slightly different purpose, we calculated the same metrics of precision and recall that we calculated for people and companies. We compared the performance of the Facenet model that was trained on VGGFace2 \cite{DBLP:conf/fgr/CaoSXPZ18} on the Labeled Faces in the Wild dataset \cite{Huang_labeledfaces}.  Our results show a recall of 80\%, and a precision of 93\%.  
\end{itemize}

### 2018-08-23 paper2018/main.tex
new version of intro

\maketitle
\begin{abstract}
AAAI creates proceedings, working notes, and technical reports directly from electronic source furnished by the authors. To ensure that all papers in the publication have a uniform appearance, authors must adhere to the following instructions. 
\end{abstract}

\input{introduction}
\input{siamese_networks}

>>>>

\maketitle
\begin{abstract}
Merging datasets is a key operation for data analytics.  A frequent requirement for merging is joining across columns that have different surface forms for the same entity (e.g., the name of a person might be represented as \textit{Douglas Adams} or \textit{Douglas Noel Adams}).  Data management systems are currently limited to performing merges based on string equality, or at best use string similarity to allow so-called `fuzzy' joins.  We propose a novel approach to performing merges inspired in part by recent advances in deep learning.  Our approach depends on (a) creating a function that maps surface forms into a set of vectors such that alternate forms for the same entity are closest in vector space, (b) indexing these vectors using a nearest neighbors algorithm to find the forms that can be potentially joined together.  Using Wikidata as our ground truth, we built models for datasets with approximately 1.1M people's names (200K identities) and 130K company names (70K identities), where the vector embeddings allow joins with precision of .91-.93 and recall of .70-.81. To build these models, we developed novel techniques for the selection of samples to train on, as well as a loss function, which we describe in the paper.  We also make the models available to the community for use in joining names or for aligning people or companies across multiple ontologies.

\end{abstract}

\input{introduction2}
\input{siamese_networks}
\input{joins}
\input{datasets}
